\normalfont \normalsize \textsc{TinyTim PSF} \\ [25pt] \horrule{0.5pt} \\[0.4cm] \huge Data Preparation\\ \horrule{2pt} \\[0.5cm]

### 2015-05-30 summary/TinyTimPSF/TinyTimPSF.tex
do tinytim psf photometry

\newcommand{\horrule}[1]{\rule{\linewidth}{#1}} % Create horizontal rule command with 1 argument of height

\title{ 
\normalfont \normalsize 
\textsc{Exoplanet Patchy Cloud Project} \\ [25pt] % Your university, school and/or department name(s)
\horrule{0.5pt} \\[0.4cm] % Thin top horizontal rule
\huge Systematic Error Correction\\ % The assignment title
\horrule{2pt} \\[0.5cm] % Thick bottom horizontal rule
}

@@

\maketitle % Print the title
\section{Summary}

According to previous results, there are two kinds of systematic errors
that downgrade the quality of photometry and need to be corrected.
\begin{enumerate}
\item images that are taken with different dithering positions have
  significant difference in photometry. The most possible cause is 
  bad flat fielding.
\item Short timescale variation on the light curves. Red noise needs
  to be caught and corrected. Following effects can contribute to red
  noise:
  ramp effect, jitter of the telescope, telescope temperature
  fluctuation, intra-pixel  response fluctuation, cosmic rays.
\end{enumerate}

  \section{Red Noises}
  \subsection{Ramp Effect and Pointing Problem}
  According to Deming et al. 2013, Wilkins 2014, ramp effect becomes significant
  when exposure level is larger than 30000 e$^{-}$ per pixel. For both
  F125W and F160W images, the largest count rate per pixel at the
  planetary companions place is 2000 - 2200 e$^{-}$s$^{-}$. For F125W
  images, the exposure time is 30s for each image while for F160W is
  15s. Therefore, the pixel that is the peak of the PSF of the
  companion object in the F125W image may be affected by ramp effect.

  In Figure \ref{fig:peak}, I plotted the count for one pixel that has
  the largest flux at the PSF region of ABPIC-B for every image. In short time-scale
  (i.e. in one exposure set, does not demonstrate in the plot), peak
  fluxes do not have a clear and common ramp effect pattern. For first
  3 orbits, the counts have large scatterings for different dithering
  positions. because the peaks of the PSF could locate at the center
  or edge of the pixel. For the last 3 orbits for which dithering is not
  applied, the peak pixel count demonstrates clear trend.  In the 4th
  and 6th orbits, the peak pixel count decreases with time while for
  5th orbits, it slightly increases with time.

    \begin{figure}[th]
    \centering
    \includegraphics[width=\textwidth]{ramp1Pix}
    \caption{Count for peak pixel on the image of ABPIC-B. Different
      color indicate different dithering position. Count rates are measured
      from the original \texttt{ima} files (no primary subtraction applied)}
    \label{fig:peak}
  \end{figure}

  First of all, I cannot be certain whether ramp effect plays a
  significant role in this trend. The exposure levels of peak pixel for
  F125W images are more than 70000 e$^{-}$, for 5th orbit they are
  near 90000 e$^{-}$, which are almost reach the fully saturation
  level of 100000 e$^{-}$. However, short timescale variations are
  random and no 'hook shaped' (Wilkins et al. 2014) light curves
  appeared.

  This trends for peak pixel count are correlated with the positions of
  PSF centroids. I plot the PSF centroids that are measured by Gaussian
  fitting for 4th, 5th and 6th orbits in Figure \ref{fig:center}. The
  PSF centroids shifted in a specific direction within one orbit. In
  4th and 6th orbits, the centroids shift to +y direction and in 5th
  orbit they shift to (+x, +y) direction. In 4th and 6th orbits, the
  centroids shift to the edge of the pixel while in 5th orbit, they
  slightly shift to the pixel center.

  \begin{figure}[th]
    \centering
    \includegraphics[width=1.0\textwidth]{Ymove}
    \caption{PSF centroid shift for ABPIC-B in 4th, 5th, and 6th
      orbits. Y-axes are the centroid coordinates $(x, y)$ minus the
      those for the first image in each orbit $(x_0, y_0)$. Red dots
      are for x-direction and blue + are for y-direction.}
    \label{fig:center}
  \end{figure}

  To sum,
  \begin{enumerate}
  \item Peak pixels have high exposure levels, especially for F125W,
    ramp effect needs to be considered.
  \item Telescope has a pointing shift within each orbit, and caused
    the peak pixel count varies in a specific trend. Since peak pixel
    contains $\sim 25\%$ of the total flux, if there were a
    considerable intra-pixel response variation, this pointing shift
    could be an error source.
  \end{enumerate}

  \subsection{Cosmic Rays in Up-the-Ramp Fitting}
  \begin{itemize}
  \item In majority of \texttt{ima} files, there are several pixels in
    companion image region that are flagged with 8192 (CR detected by
    up-the-ramp fitting)
    \item it is not clear yet number of pixels that have CR flag is
      correlated with flux scattering
    \item considering use the same PSF images for PSF subtraction.
    \end{itemize}

    \subsection{Jitter}
    \begin{itemize}
    \item 4 exposures have extremely large jitter value
     \item RMS jitter values are about 0.001 arcsec (0.1 pixel)in both
       axes
     \item jitter values seem not to correlate with flux scattering
    \end{itemize}
\end{document}
%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:

>>>>

\newcommand{\horrule}[1]{\rule{\linewidth}{#1}} % Create horizontal rule command with 1 argument of height
\newcommand{\chisq}{\ensuremath{\chi^2}}
\title{ 
\normalfont \normalsize 
\textsc{TinyTim PSF} \\ [25pt] % Your university, school and/or department name(s)
\horrule{0.5pt} \\[0.4cm] % Thin top horizontal rule
\huge Data Preparation\\ % The assignment title
\horrule{2pt} \\[0.5cm] % Thick bottom horizontal rule
}

@@

\maketitle % Print the title

This document is a summary of using TinyTim PSF to measure the
photometry of 2M1207 system.

2M1207 system has a very small separation between the primary and
secondary. The peak of primary and secondary separates about 6
pixels. A perfect match of primary image and PSF is required at an
extreme level to ensure a good photometry measurement of
2M1207b. Image of the same system with different roll angle is an
option for PSF subtraction. However, considering coarse sample rate of
WFC3 IR, to shift the image can generate large artifact no matter what
interpolation method is using. Also, in our data, the angle telescope
rolled is ~30 degree, which could cause self-subtraction of the
secondary.

The advantage of TinyTim PSF is that TinyTim can produce 10x super
sampled PSF so that image shifting and interpolation is no longer a
problem. The downside of TinyTim is that it has several knobs, such as
jittering, secondary mirror shift to tune and even using the best
matched parameter set, TinyTim PSF are not good at modeling
diffraction spiders and coma due to intrinsic problems.

\section{Generate PSF}

\begin{enumerate}
  \item specify spectrum for PSF using \texttt{tiny1}, and generate
    the input script for \texttt{tiny2}
  \item using python script \texttt{PSF\_generator.py} to modify the
    input script and run \texttt{tiny2} and \texttt{tiny3} to generate
    PSF\\
    Lines in the input script that need to be changed:

    \begin{itemize}
    \item line 2, the output file name
    \item line 8, the filter
    \item line10, major axis jitter in mas
    \item line 11, minor axis jitter in mas
    \item line 14, the x and y position on the detector, in integer
    \item line 260, secondary mirror displacement
    \end{itemize}
    \item the way to generate a 10x sampled PSF using \texttt{tiny3}:
      \texttt{tiny3 input.script sub=10}
\end{enumerate}

\section{Least \chisq Fit}

To match the center of PSF and star image, fit the PSF to image in a
grid of $dx$ and $dy$. The least \chisq fit calculates the amplitude
of PSF and a sky level. Amplitude and sky level can be solved
analytically.
\newcommand{\PSF}{\ensuremath{\mathrm{PSF}}}
\newcommand{\Img}{\ensuremath{\mathrm{Img}}}
\begin{equation}
  \begin{split}
    \chisq &= \sum_{i} \frac{(\mathrm{Img}_{i} - a\cdot\mathrm{PSF}_{i} -
      b)^{2}}{\sigma_{i}^{2}}\\
    &= \sum_{i} \frac{\mathrm{Img}_{i}^{2}+a^{2}\cdot \mathrm{PSF}_{i}^{2}+b^{2}+2ab\cdot \mathrm{PSF}_{i}
      -2a\cdot \mathrm{Img}_{i}\PSF_{i}-2b \mathrm{Img}_{i}}{\sigma_{i}^{2}}
  \end{split}
    \label{eq:chisq}
\end{equation}
To obtain least \chisq, $\frac{\partial \chisq}{\partial a} = 0$,
$\frac{\partial \chisq}{\partial b}=0$. So that
\begin{equation}
  \label{eq:calchisq}
  \begin{bmatrix}
    &\sum_{i}\frac{\PSF_{i}^{2}}{\sigma_{i}^{2}}
    &\sum_{i}\frac{\PSF_{i}}{\sigma_{i}^{2}}\\
    &\sum_{i}\frac{\PSF_{i}}{\sigma_{i}^{2}}
    &\sum_{i}\frac{b}{\sigma_{i}^{2}}
  \end{bmatrix}
  \cdot
  \begin{bmatrix}
    &a\\
    &b
  \end{bmatrix}
  =
  \begin{bmatrix}
    &\sum_{i}\frac{\Img_{i}\cdot\PSF_{i}}{\sigma_{i}^{2}}\\
    &\sum_{i}\frac{\Img_{i}}{\sigma_{i}^{2}}
  \end{bmatrix}
\end{equation}
$a$ and $b$ can be easily solved using simple linear algebra.

\section{Fit 2 PSFs together.}
Simply change equation \ref{eq:calchisq} to
\newcommand{\PSFI}{\ensuremath{\mathrm{PSF1}}}
\newcommand{\PSFII}{\ensuremath{\mathrm{PSF2}}}
\begin{equation}
  \label{eq:1}  
  \begin{bmatrix}
    & \sum_{i} \frac{\PSFI_i^2}{\sigma^2}
    &\sum_{i}\frac{\PSFI_i\cdot\PSFII_i}{\sigma_i^2}
    &\sum_{i}\frac{\PSFI_i}{\sigma_i^{2}}\\
    & \sum_{i} \frac{\PSFI_i\cdot\PSFII_{i}}{\sigma^2}
    &\sum_{i}\frac{\PSFII_i^{2}}{\sigma_i^2}
    &\sum_{i}\frac{\PSFII_i}{\sigma_i^{2}} \\
    &\sum_{i}\frac{\PSFI_{i} }{\sigma_{i}^{2}}
    &\sum_{i}\frac{\PSFII_{i} }{\sigma_{i}^{2}}
    &\sum_{i}\frac{b}{\sigma_{i}^{2}}
  \end{bmatrix}
  \cdot
  \begin{bmatrix}
    &a_1\\
    &a_2\\
    & b
  \end{bmatrix}
  =
  \begin{bmatrix}
    &\sum_{i}\frac{\Img_i\PSFI_i}{\sigma_i^2}\\
    &\sum_i \frac{\Img_i\PSFII_i}{\sigma_i^2}\\
    &\sum_{i}\frac{\Img_{i}}{\sigma_{i}^{2}}
  \end{bmatrix}
\end{equation}
\end{document}
%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:
